---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
    fig_height: 4
    fig_width: 6
    number_sections: yes
    theme: readable
---


# Loading and preprocessing the data 
    The first step consists in loading the data, getting rid of the N.A value, and coercing the date by `as.DATE`
```{r loading data, echo=TRUE, tidy=TRUE}
    setwd("C:/Users/trand_000/Desktop/Coursera Data/reproductible research/RepData_PeerAssessment1")
    activityOld<- read.csv("activity.csv")
    activityOld$date<- as.Date(activityOld$date, "%Y-%m- %d")
    activity<- activityOld[!is.na(activityOld$step),]
    head(activity)

```
# What is mean, median and total number of steps taken per day?
##  Calculate total number of steps per day
```{r sum, echo= TRUE}
    sum<- tapply(activity$step, activity$date, sum)
```
## Histogram    
``` {r barplot, fig.height=4}
    barplot(height=sum, names=("")) 
    title("TOTAL STEPS PER DAY")
    box()
    axis(1, at=c(1,26,53), lab=c("02 Oct","28 Oct","29 Nov"))
```
   
## Mean and median
```{r mean, echo= TRUE}
    mean<-mean(sum) 
    median<- median(sum)
```
    The mean of total step per day is `r mean`. The median is `r median`
# What is the average daily activity pattern?
## Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
    Calculate the pattern by interval
```{r pattern}
    patt <- aggregate(activity$step ~ activity$interval, 
                  data=activity, FUN="mean")
    head(patt)
```
    The time seris plot
```{r lineplot, fig.height=4}
    plot(patt, type="l", main="Average daily activity pattern", 
        xlab="Interval", ylab="Steps")
    lines(patt, col="blue")
```

## Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r max, echo=TRUE}
    max<-match(max(patt[,2]), patt[,2])
    max<-patt[max,1]
``` 
    The 5-minute interval containing the maximu number of stap is `r max`
# Imputing missing values
  Note that there are a number of days/intervals where there are missing values (coded as NA). The presence of missing days may introduce bias into some calculations or summaries of the data.
##Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r mis_Val, echo=TRUE}
    u<- is.na(activityOld)
    na<-sum(u)
```
    The total number of missing values in the dataset is `r na`
## Filling the Dataset
    I'll fill the missing value by the mean for that 5-minute interval
```{r fill, echo=TRUE}
    activityNew<- activityOld
    aves <- aggregate(activityNew$step ~ activityNew$interval, 
                  data=activityNew, FUN="mean")
    for (i in 1:nrow(activityNew)){
        if (is.na(activityNew[i, 1])) {
            position<-match(activityNew[i, 3], aves[,1])
            activityNew[i,1]<- aves[position, 2]
        }
    }
    head(activityNew)
```
## Histogram of the total number of steps taken each day 

```{r barplot2,fig.height=4}
    sumNew<- tapply(activityNew$step, activityNew$date, sum)
    barplot(height=sumNew, main="Total step per day (imputing NA)", xlab="Date", ylab="Steps")
    box()

```
## Calculate and report the mean and median total number of steps taken per day.
```{r meanNew, echo=TRUE}
    meanNew<- mean(sumNew) 
    medianNew<- median(sumNew)
    print(meanNew)
    print(medianNew)
```
    The 'old' mean and 'old' median are `r mean`, `r median`. There are no huge difference between these values 
# Are there differences in activity patterns between weekdays and weekends?
## "Weekday" or "Weekend"?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r weekend}
    library(lubridate)
    date<- wday(activityNew$date)
    u<- date==1|date==7
    activityNew["weekend"]<-u
```
## Plot
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r lineplot2}
    act1<- activityNew[!u,]
    act2<- activityNew[u,]
    pattWday<- tapply(act1$step, act1$interval, mean)
    pattWend<- tapply(act2$step, act2$interval, mean)
    ##Create multiple plot##
    par(mfrow=c(2,1))
    


```